---
title: "Statistical Computation HW1"
author: "Jennifer Lieu"
date: "9/15/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Part 1: Importing Data into R

i.
```{r}
#set working directory and importing data
setwd("/Users/jenniferlieu/Desktop/Stat_Comp_Data")
titanic <- read.table("Titanic.txt", header=FALSE, as.is=TRUE)
```

ii.
```{r}
#getting summary of data
str(titanic)
```
From observing the output of the R code, we can tell that there are 12 variables, so 12 columns, and 892 obsercations, so 891 rows because we have the first row being the labels.

iii.
```{r}
# initialize vector
v13 <- c(1,2,3)
# Set label
v13[1] <- "Survived.Word"
#initialize parameter
i=2
#loop the 'survived' vector, and set v13 to survived or died depending on values in 'survived' vector.
while(i < 893){
  if(titanic[i,2] =="1"){
    v13[i]="survived"
  }
  else{
    v13[i]="died"
  }
  i=i+1
}
#adding column to dataset
titanic$v13 <- v13
```

Part 2:Exploring the data

i. 
```{r}
c <- c(2,6,10)
datamatrix <- data.matrix(titanic[2:892,c])
apply(datamatrix, 2, mean)
```
The resulting mean for survived is the fraction of passengers who survived the titanic. The reason why Age has a mean of 'NA' is because we do not have all the age values for NA in our dataset, and therefore we cannot calculate the average Age of all the passengers on the voyage. 

ii.
```{r}
titanicf <- titanic[titanic$V5 == "female",]
datamatrixf <- data.matrix(titanicf[1:314,2])
datamatrixf <- as.numeric(datamatrixf)
round(sum(datamatrixf)/891,2)
```
The proportion of female passengers who survived the titanic disaster is .26 when rounding to the nearest 2 digits.

iii.
```{r}
titanics <- titanic[titanic$v13 == "survived",]
round(length(titanics$V5[titanics$V5 == "female"])/(length(titanics$v5)))
round(length(titanics$V5[titanics$V5 == "female"])/length(titanics$V5),2)
```
Out of all the survivors, .68 of them are females when rounding to the nearest 2 digits.

iv. 
```{r}
classes <- as.numeric(sort(unique(titanic$V3)[2:892]))
Pclass.Survival <- vector("numeric", length = 3)
names(Pclass.Survival) <- classes

for(i in 1:3){
  # filter the classes
  titanicp <- titanic[titanic$V3 == as.character(i),]
  # filter the survived
  titanicps <- titanicp[titanicp$v13 == "survived",]
  # length of vector after filters
  numerator <- length(titanicps$V3)
  # length of vector after filtering only classes
  denominator <- length(titanic$V3[titanic$V3 == as.character(i)])
  # fraction
  fraction <- numerator/denominator
  #add fraction to the vector
  Pclass.Survival[i] <- round(fraction,2)
}
```

v. 
```{r}
classes2 <- as.numeric(sort(unique(titanic$V3)[2:892]))
Pclass.Survival2 <- vector("numeric", length = 3)
names(Pclass.Survival2) <- classes2
factor2 <- factor(titanic$V3[2:892])

fraction <- function(i){
   proportion <- round(length(i[i == "survived"])/length(i),2)
   return(proportion)
}

Pclass.Survival2 <- tapply(titanic$v13[2:892], factor2, fraction)
```

vi.
As the class number gets lower, the survival rate gets higher.